---
title: "tils_visualisation"
author: "Michael Monsonego"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    fig_crop: no
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width=17) 
```
## 1. number of cells per patient + tSNE plot
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height= 12}
library(dplyr)
library(Seurat)
library(patchwork)
library(tidyverse)
library(sleepwalk)
library(gridExtra)
library(future)
library(reshape2)
library(reticulate)
library(ggplot2)
library(plyr)
library(stringr)
library(RColorBrewer)
library(grid)
library(cowplot)
library(ggsignif)
library(DT)
library(Cairo)
library(snakecase)
library(glue)

setwd("D:/Michael/git_check/tils")
x1=4.6
T_cells = readRDS("objects/tils_all_.45_integrgate.rds")

# counting cells in treatment and control
unique(T_cells$orig.ident) # R1-4, N1-4
table(T_cells$orig.ident)["R1"] #  
table(T_cells$orig.ident)["R2"] #  
table(T_cells$orig.ident)["R3"] #  
table(T_cells$orig.ident)["R4"] #  
table(T_cells$orig.ident)["N1"] #  
table(T_cells$orig.ident)["N2"] #  
table(T_cells$orig.ident)["N3"] #  
table(T_cells$orig.ident)["N4"] #  

T_cells  = RunTSNE(T_cells , dims = 1:15, verbose = F)
DimPlot(T_cells , reduction = "tsne", label= TRUE, pt.size=0.5, label.size = 2)
```

Top DE genes
--------------
```{r echo=FALSE, cashe = TRUE}
Top50Markers <- read.delim("D:/Michael/git_check/tils/excels/Top50Markers_perClust.tils.45integrgate.csv", sep = ",", header = T, quote="\"", check.names=FALSE)
Top50Markers$avg_log2FC <- Top50Markers$avg_log2FC %>% round(2)
Top50Markers <- Top50Markers[,c("cluster","gene", "avg_log2FC", "p_val")] 
datatable(Top50Markers, rownames = FALSE, filter = 'top', colnames = c('Cluster #', 'Gene Name', "log2(FC)", 'P.value'))
```

## Contribution of each group 
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= 10}
#M# unnormalized counts per cluster per treatment.
ggplot(T_cells@meta.data, aes(x=T_cells@active.ident, fill=Treatment)) + geom_bar(stat="count",position=position_dodge())+
    scale_fill_manual(values= c("#CFBAE1", "#5C7D9D"))+
    theme_classic()+
    labs(x="")+
    theme(legend.title=element_text(size=18),legend.text= element_text(size=14), 
          axis.text.x=element_text(size=12,angle=45, hjust=1))+labs(title = "unnormalized counts per cluster per treatment")

ggsave(file = "figures/Tcells/unnormalized_counts_per_cluster_per_treatment.png", dpi=300, width=15, height=10)
```

```{r message=FALSE, warning=FALSE, cache= TRUE,  fig.height= 6}
#M# percentage of contribution to each cluster per treatment
ggplot(T_cells@meta.data, aes(x=T_cells@active.ident, fill=Treatment)) + geom_bar(position="dodge")+
    scale_fill_manual(values= c("#CFBAE1", "#5C7D9D"))+
    theme_classic()+
    labs(x="")+
    theme(legend.title=element_text(size=18),legend.text= element_text(size=14), 
          axis.text.x=element_text(size=12,angle=45, hjust=1))+labs(title = "percentage of contribution to each cluster per treatment")
ggsave(file = "figures/Tcells/percentage_of_contribution_stacked_bar_plot.png", dpi=300, width=10, height=10)
```

## unnormalized contribution per patient 
```{r message=FALSE, warning=FALSE, cache= TRUE,  fig.height= 6}
Sample_Names_vec <- c("N1","N2","N3","N4")
for (sample in Sample_Names_vec) {
  Tsub = T_cells@meta.data[T_cells@meta.data$orig.ident == sample, ]
 ggplot(Tsub, aes(x=Tsub$seurat_clusters, fill=Treatment)) + geom_bar(stat="count",position=position_dodge())+
    scale_fill_manual(values= c("#CFBAE1"))+
    theme_classic()+
    labs(x="")+
    theme(legend.title=element_text(size=18),legend.text= element_text(size=14), 
          axis.text.x=element_text(size=12,angle=45, hjust=1))+labs(title = glue({sample}, " : unnormalized counts per cluster per treatment"))
  ggsave(filename = glue({sample}, "_unnormalized_counts_per_cluster_per_treatment.png"), path = "figures/Tcells/", dpi=300, width=10, height=10)
}

Sample_Names_vec <- c("R1","R2","R3","R4")
for (sample in Sample_Names_vec) {
  Tsub = T_cells@meta.data[T_cells@meta.data$orig.ident == sample, ]
 ggplot(Tsub, aes(x=Tsub$seurat_clusters, fill=Treatment)) + geom_bar(stat="count",position=position_dodge())+
    scale_fill_manual(values= c("#5C7D9D"))+
    theme_classic()+
    labs(x="")+
    theme(legend.title=element_text(size=18),legend.text= element_text(size=14), 
          axis.text.x=element_text(size=12,angle=45, hjust=1))+labs(title = glue({sample}, " : unnormalized counts per cluster per treatment"))
  ggsave(filename = glue({sample}, "_unnormalized_counts_per_cluster_per_treatment.png"), path = "figures/Tcells/", dpi=300, width=10, height=10)
}
```
## normalized contribution per patient 
```{r echo=FALSE, fig.height=4, fig.width=10}
frque <- table(Idents(T_cells), T_cells$Treatment)
frque <- as.data.frame(frque)
frque <- frque %>% dplyr::rename(Cluster = Var1, Treatment = Var2)
frque$Treatment <- factor(frque$Treatment)

frque <- ddply(frque, .(Treatment),  transform, percentperstatus   = Freq/sum(Freq))
frque <- ddply(frque, .(Cluster), transform, percentpercluster  = Freq/sum(Freq))
frque <- ddply(frque, .(Cluster), transform, normalizedppstatus = percentperstatus/sum(percentperstatus)*100)
frque <- ddply(frque, .(Treatment),  transform, normalizedpcluster = percentpercluster/sum(percentpercluster))

mfrque <- table(Idents(T_cells), T_cells$orig.ident)
mfrque <- as.data.frame(mfrque)
mfrque <- mfrque %>% dplyr::rename(Cluster = Var1, patient = Var2)

mfrque <- ddply(mfrque, .(patient),  transform, percentperstatus   = Freq/sum(Freq))
mfrque <- ddply(mfrque, .(Cluster), transform, percentpercluster  = Freq/sum(Freq))
mfrque <- ddply(mfrque, .(Cluster), transform, normalizedppstatus = percentperstatus/sum(percentperstatus)*100)
mfrque <- ddply(mfrque, .(Treatment),  transform, normalizedpcluster = percentpercluster/sum(percentpercluster))





p3 <- ggplot(mfrque, aes(x = Cluster, y = normalizedppstatus, fill = patient)) +
 theme_minimal() +
 geom_bar(stat = "identity", width = 0.7,position="dodge") +
   scale_fill_manual(values= c("#CFBAE1", "#A8A5CA", "#8291B3", "#5C7D9D", "#FF6F61", "#D78459", "#AF9A52", "#88B04B"))+
 ggtitle("Normalized Cluster Counts") +
 ylab("Normalized Proportions in %") +
 theme(
   axis.text.x  = element_text(size = 11, color = "black"),
   axis.text.y  = element_text(size = 11, color = "black"),
   axis.title.x = element_text(size = 13, color = "black"),
   axis.title.y = element_text(size = 13, color = "black"),
   title = element_text(size = 13),
   legend.justification = "top",
   panel.border = element_blank()) +
 geom_hline(yintercept = 20, linetype='dotted')
  
  
p1 <- ggplot(frque, aes(x = Cluster, y = normalizedppstatus, fill = Treatment)) +
 theme_minimal() +
 geom_bar(stat = "identity", width = 0.7,position="dodge") +
 scale_fill_manual(values= c("#CFBAE1", "#5C7D9D"))+
 ggtitle("Normalized Cluster Counts") +
 ylab("Normalized Proportions in %") +
 theme(
   axis.text.x  = element_text(size = 11, color = "black"),
   axis.text.y  = element_text(size = 11, color = "black"),
   axis.title.x = element_text(size = 13, color = "black"),
   axis.title.y = element_text(size = 13, color = "black"),
   title = element_text(size = 13),
   legend.justification = "top",
   panel.border = element_blank()) +
 geom_hline(yintercept = 33.3, linetype='dotted') 

p2<-ggplot(frque, aes(x= Cluster, y= Freq, fill = Treatment)) +
 theme_minimal() +
 geom_bar(stat = "identity", width = 0.7) +
 scale_fill_manual(values= c("#CFBAE1", "#5C7D9D"))+
 ggtitle("Cluster Counts") + ylab("Cell counts\n") +
 theme(
   axis.text.x  = element_text(size = 11, color = "black"),
   axis.text.y  = element_text(size = 11, color = "black"),
   axis.title.x = element_text(size = 13, color = "black"),
   axis.title.y = element_text(size = 13, color = "black"),
   title = element_text(size = 13),
   legend.justification = "top",
   panel.border = element_blank())

p2+p1+plot_layout(guides = "collect")+plot_annotation(tag_levels = 'A') 
p3
```



Markers
=====

### Proliferation Markers

```{r cache=T, fig.height=x1*1, echo=FALSE}
Prolif <- c("TOP2A","MKI67")
FeaturePlot(T_cells, features = Prolif, label.size = 8, pt.size = 1, label=T, ncol = 3, order = T,reduction = "tsne")
```

### T Cells Markers
```{r cache=T, fig.height=x1*2,warning=FALSE, echo=FALSE}
Ts <- c('Cd4', 'Cd8a', 'Cd3d','Cd3e','Cd3g',"PTPRC")
Ts <- toupper(Ts)
FeaturePlot(T_cells, features = Ts, label.size = 8, pt.size = 1, label=T, ncol = 3, order = T,reduction = "tsne")
```


### a. activation/exhaustion markers
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= 16}
# activation / exhaustion markers
FeaturePlot(T_cells, features = c("PDCD1","HAVCR2","LAG3","TIGIT","TNFRSF4","TNFRSF9","ICOS","CTLA4","KLRD1","IRF8","KLRC1","TOX"), order=TRUE,pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_activation.png", dpi=300, width=10, height=10)


upper_vector <- toupper(c("PDCD1","HAVCR2","LAG3","TIGIT","TNFRSF4","TNFRSF9","ICOS","CTLA4","KLRD1","IRF8","KLRC1","TOX"))
quoted_vector <- sprintf('"%s"', upper_vector)
upper_string_with_quotes <- paste(quoted_vector, collapse = ",")
cat(upper_string_with_quotes)
```
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=16}
VlnPlot(T_cells, features = c("PDCD1","HAVCR2","LAG3","TIGIT","TNFRSF4","TNFRSF9"), stack=TRUE, flip= TRUE, split.by = "Treatment")+theme(axis.text.y =element_text(size=6))+ theme_classic()+theme(axis.text.x = element_text(angle = 70, hjust=1, size=12), axis.title.x = element_blank()) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/vln_activation_exhastion_cluster_by_treatment_1.png", dpi=300, width=8, height=10)

VlnPlot(T_cells, features = c("PDCD1","HAVCR2","LAG3","TIGIT","TNFRSF4","TNFRSF9"), stack=TRUE, flip= TRUE, group.by = "Treatment", fill.by ="ident")+theme(axis.text.y =element_text(size=6))+ theme_classic()+theme(axis.text.x = element_text(angle = 70, hjust=1, size=12), axis.title.x = element_blank()) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/vln_activation_exhastion_by_treatment_1.png", dpi=300, width=4, height=10)

VlnPlot(T_cells, features = c("PDCD1","HAVCR2","LAG3","TIGIT","TNFRSF4","TNFRSF9"), stack=TRUE, flip= TRUE, fill.by = "ident")+theme(axis.text.y =element_text(size=6))+ theme_classic()+theme(axis.text.x = element_text(angle = 70, hjust=1, size=12), axis.title.x = element_blank()) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/vln_activation_exhastion_by_cluster_1.png", dpi=300, width=6, height=10)

##

VlnPlot(T_cells, features = c("ICOS","CTLA4","KLRD1","IRF8","KLRC1","TOX"), stack=TRUE, flip= TRUE, split.by = "Treatment")+theme(axis.text.y =element_text(size=4))+ theme_classic()+theme(axis.text.x = element_text(angle = 70, hjust=1, size=12), axis.title.x = element_blank()) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_activation_exhaustion_cluster_by_treatment_2.png", dpi=300, width=8, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("ICOS","CTLA4","KLRD1","IRF8","KLRC1","TOX"), stack=TRUE, flip= TRUE, group.by = "Treatment", fill.by ="ident")+theme(axis.text.y =element_text(size=4))+ theme_classic()+theme(axis.text.x = element_text(angle = 70, hjust=1, size=12), axis.title.x = element_blank()) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_activation_exhaustion_by_treatment_2.png", dpi=300, width=4, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("ICOS","CTLA4","KLRD1","IRF8","KLRC1","TOX"), stack=TRUE, flip= TRUE, fill.by = "ident")+theme(axis.text.y =element_text(size=4))+ theme_classic()+theme(axis.text.x = element_text(angle = 70, hjust=1, size=12), axis.title.x = element_blank()) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_activation_exhaustion_by_cluster_2.png", dpi=300, width=6, height=10, limitsize=FALSE)

```

### effector markers 
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=16}
FeaturePlot(T_cells, features = c("GZMB","PRF1","IFNG","MKI67","GZMC","GZMA","KLRG1","IL7R","CD52"),order=TRUE,pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_effector.png", dpi=300, width=8, height=10)
```
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=16}
VlnPlot(T_cells, features = c("GZMB","PRF1","IFNG"), stack=TRUE, flip= TRUE, split.by = "Treatment")+theme(axis.text.y =element_text(size=6))+ theme_classic()+theme(axis.text.x = element_text(angle = 70, hjust=1, size=12), axis.title.x = element_blank()) + theme(axis.title.y=element_text(size=12)) +
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_effector_cluster_by_treatment_1.png", dpi=300, width=8, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("GZMB","PRF1","IFNG"), stack=TRUE, flip= TRUE, group.by = "Treatment", fill.by ="ident")+theme(axis.text.y =element_text(size=6))+ theme_classic()+theme(axis.text.x = element_text(angle = 70, hjust=1, size=12), axis.title.x = element_blank()) + theme(axis.title.y=element_text(size=12)) +
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_effector_by_treatment_1.png", dpi=300, width=4, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("GZMB","PRF1","IFNG"), stack=TRUE, flip= TRUE, fill.by = "ident")+theme(axis.text.y =element_text(size=6))+ theme_classic()+theme(axis.text.x = element_text(angle = 70, hjust=1, size=12), axis.title.x = element_blank()) + theme(axis.title.y=element_text(size=12)) +
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_effector_by_cluster_1.png", dpi=300, width=6, height=10, limitsize=FALSE)

##

VlnPlot(T_cells, features = c("GZMC","GZMA","KLRG1","IL7R","CD52"), stack=TRUE, flip= TRUE, split.by = "Treatment")+theme(axis.text.y =element_text(size=6))+ theme_classic()+theme(axis.text.x = element_text(angle = 70, hjust=1, size=12), axis.title.x = element_blank()) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_effector_cluster_by_treatment_2.png", dpi=300, width=8, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("GZMC","GZMA","KLRG1","IL7R","CD52"), stack=TRUE, flip= TRUE, group.by = "Treatment", fill.by ="ident")+theme(axis.text.y =element_text(size=6))+ theme_classic()+theme(axis.text.x = element_text(angle = 70, hjust=1, size=12), axis.title.x = element_blank()) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_effector_by_treatment_2.png", dpi=300, width=4, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("GZMC","GZMA","KLRG1","IL7R","CD52", fill.by = "ident"), stack=TRUE, flip= TRUE)+theme(axis.text.y =element_text(size=6))+ theme_classic()+theme(axis.text.x = element_text(angle = 70, hjust=1, size=12), axis.title.x = element_blank()) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_effector_by_cluster_2.png", dpi=300, width=6, height=10, limitsize=FALSE)

```


### c. naive-like (naive or memory cells)
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= 10}
# naive
FeaturePlot(T_cells, features = c("SELL","IL7R","CCR7","MS4A4C"),order=TRUE,pt.size=1, reduction="tsne", ncol=2)
ggsave(file = "figures/Tcells/features_T_naive.png", dpi=300, width=8, height=10)
```

```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= 10}
# memory 
FeaturePlot(T_cells, features = c("SELL","CD44","IL7R","CCR7","TCF7","CXCR6"),order=TRUE,pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_memory.png", dpi=300, width=10, height=8)
```

### d. effector memory markers
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= x1*4}
FeaturePlot(T_cells, features = c("BACH2","PRDM1","IL7R","CD44","CX3CR1","ID2","KLF2","KLRG1","LEF1","STAT4","TBX21","ZEB2"),order=TRUE,pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_effector memory.png", dpi=300, width=30, height=40)
```

### e. central memory markers
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= x1*5}
FeaturePlot(T_cells, features = c("TCF7","BACH2","BCL16","CCR7","IL7R","CD27","CD44","SELL","ID3","KLF2","LCR7","LEF1","SLAMF6","STAT3","TFAP4"),order=TRUE,pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_central memory.png", dpi=300, width=30, height=50, limitsize=FALSE)
```


### f. resident memory markers
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= x1*8}
FeaturePlot(T_cells, features = c("PRDM1","ITGAE","IL7R","CD44","ITGA1","CD69","CRTAM","DUSP6","EGR2","FOSB","FOXO3","GZMA","GZMB","ZNF683","ID3","IL2RA","IRF4","JUN","JUNB","MYC","NR4A1","RGS1"), order=TRUE, pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_resident memory.png", dpi=300, width=30, height=80, limitsize=FALSE)
```

### g. B cells
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=x1*3}
FeaturePlot(T_cells, features = c("CD19","MS4A1","IGLC1","IGHM","IGKC","CD3E","CD8A"), order=TRUE,pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/Features_T_B + CD8 cells.png", dpi=300, width=30, height=30)
```

## 3. Heatmap 

### a. activation/exhaustion/effector
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=12}
DoHeatmap(T_cells, features = c("PDCD1","HAVCR2","LAG3","TIGIT","TNFRSF4","TNFRSF9","ICOS","CTLA4","KLRD1","IRF8","KLRC1","TOX1","CX3CR1","XCL1","GZMB","PRF1","MKI67","IFNG","GZMC","GZMA","KLRG1","IL7R","CD52"))+ NoLegend() + scale_fill_gradientn(colors = c("blue", "white", "red"))
ggsave(file = "figures/Tcells/T_heatmap_activated effector.png", dpi=300, width=30, height=30)
```

### b. naive/memory
```{r fig.height=12, message=FALSE, warning=FALSE, cache=TRUE}
DoHeatmap(T_cells, features = c("SELL","IL7R","CCR7","MS4A4C","SELL","CD44","IL7R","CCR7","TCF7","CXCR6","TXK","S1PR1","LEF1","SATB1" ))+ NoLegend() + scale_fill_gradientn(colors = c("blue", "white", "red"))
ggsave(file = "figures/Tcells/T_heatmap_naive memory.png", dpi=300, width=30, height=30)
```

### c. effector memory
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=12}
DoHeatmap(T_cells, features = c("BACH2","PRDM1","IL7R","CD44","CX3CR1","ID2","KLF2","KLRG1","LEF1","STAT4","TBX21","ZEB2","GZMA","GZMK","NKG7","CD8A","CD8B1","CTSW","GZMB","CCL5","CST7","PRF1","ABI3","FASL","ITM2C","1500009L16RIK","EOMES","CHST12","CCR5","HCST","AOAH","HOPX","SLAMF7","CXCR3","OASL1","F2R","CXCR6"))+ NoLegend() + scale_fill_gradientn(colors = c("blue", "white", "red"))
ggsave(file = "figures/Tcells/T_heatmap_effector memory.png", dpi=300, width=30, height=30)
```

### d. central memory
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=12}
DoHeatmap(T_cells, features = c("BACH2","BCL16","CCR7","IL7R","CD27","CD44","SELL","ID3","KLF2","LCR7","LEF1","SLAMF6","STAT3","TCF7","TFAP4"))+ NoLegend() + scale_fill_gradientn(colors = c("blue", "white", "red"))
ggsave(file = "figures/Tcells/T_heatmap_central memory.png", dpi=300, width=30, height=30)
```

### e. resident memory
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=12}
DoHeatmap(T_cells, features = c("PRDM1","ITGAE","IL7R","CD44","ITGA1","CD69","CRTAM","CX3CR1","DUSP6","EGR2","FABP5","FOS","FOSB","FOXO3","GZMA","GZMB","ZNF683","ID3","IL2RA","IRF4","JUN","JUNB","MYC","NR4A1","RGS1"))+ NoLegend() + scale_fill_gradientn(colors = c("blue", "white", "red"))
ggsave(file = "figures/Tcells/T_heatmap_resident memory.png", dpi=300, width=30, height=30)
```

### h. T cells features signatures 
```{r echo=FALSE, fig.height=12, message=FALSE, warning=FALSE, cache=FALSE}
CD8_memory_list = list(c("IL7R","CD44","CXCR6"))
T_cells = AddModuleScore(object = T_cells, features = CD8_memory_list, name = "CD8_memory", assay = "RNA")
a <- FeaturePlot(object = T_cells, features = "CD8_memory1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "0_ CD8+ memory", subtitle = "(Cxcl10, Vcan, Il18,Plcb1,Cacna1d)")+ theme(plot.subtitle = element_text(hjust = 0.5))
a
ggsave(file = "figures/Tcells/CD8_memory_Cxcl10.png", dpi=300, width=5, height=5)



activated_list = list(c("PDCD1","KLRD1","GZMA"))
T_cells = AddModuleScore(object = T_cells, features = activated_list, name = "activated", assay = "RNA")
b <- FeaturePlot(object = T_cells, features = "activated1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "activated_CD8+activated", subtitle = "(Cd4, Foxp3,Il2ra, Icos)")+ theme(plot.subtitle = element_text(hjust =0.5))
b
ggsave(file = "figures/Tcells/activated.png", dpi=300, width=5, height=5)

effector_memory_list = list(c("ID2","STAT3","GZMB","PRF1","IL7R"))
T_cells = AddModuleScore(object = T_cells, features = effector_memory_list, name = "effector_memory", assay = "RNA")
c <- FeaturePlot(object = T_cells, features = "effector_memory1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "2_ CD8+ effector memory", subtitle = "(Cd4, Foxp3,Il2ra, Icos)")+ theme(plot.subtitle = element_text(hjust =0.5))
c
ggsave(file = "figures/Tcells/effector_memory_Tregs.png", dpi=300, width=5, height=5)

cytotoxic_list = list(c("GZMB","GZMC","GRZMF"))
T_cells = AddModuleScore(object = T_cells, features = cytotoxic_list, name = "cytotoxic", assay = "RNA")
d <- FeaturePlot(object = T_cells, features = "cytotoxic1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "3_ CD8+ cytotoxic", subtitle = "(Cd4, Foxp3,Il2ra, Icos)")+ theme(plot.subtitle = element_text(hjust =0.5))
d
ggsave(file = "figures/Tcells/cytotoxic_Tregs.png", dpi=300, width=5, height=5)


CD8_naive_memory_list = list(c("CD8A","CD4","SELL","CCR7","TCF7","LEF1","CD8","ITGAE","CXCR3","GZMA","SELL","CCR7"))
T_cells = AddModuleScore(object = T_cells, features = CD8_naive_memory_list, name = "CD8_naive_memory", assay = "RNA")
e <- FeaturePlot(object = T_cells, features = "CD8_naive_memory1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "4_ CD8+ naive/memory", subtitle = "(Cd4, Foxp3,Il2ra, Icos)")+ theme(plot.subtitle = element_text(hjust =0.5))
e
ggsave(file = "figures/Tcells/CD8_naive_memory_Tregs.png", dpi=300, width=5, height=5)


CD4_CD8_naive_memory_list = list(c("CD8A","CD4","SELL","CCR7","TCF7","LEF1"))
T_cells = AddModuleScore(object = T_cells, features = CD4_CD8_naive_memory_list, name = "CD4_CD8_naive_memory", assay = "RNA")
f <- FeaturePlot(object = T_cells, features = "CD4_CD8_naive_memory1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "5_ CD4+ CD8+ naive/memory", subtitle = "(Cd8a, Cd4, Sell, Ccr7, Tcf7, Lef1)")+ theme(plot.subtitle = element_text(hjust =0.5))
f
ggsave(file = "figures/Tcells/CD4_CD8_naive_memory_CD4_CD8_naive_memory.png", dpi=300, width=5, height=5)

proliferating_list = list(c("CD8","MKI67"))
T_cells = AddModuleScore(object = T_cells, features = proliferating_list, name = "proliferating", assay = "RNA")
g <- FeaturePlot(object = T_cells, features = "proliferating1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "6_ CD8+ proliferating", subtitle = "(Cd8, Mki67)")+ theme(plot.subtitle = element_text(hjust =0.5))
g
ggsave(file = "figures/Tcells/proliferating_CD8+proliferating.png", dpi=300, width=5, height=5)

Tregs_list = list(c("CD4","FOXP3","IL2RA","ICOS","TNFRSF4"))
T_cells = AddModuleScore(object = T_cells, features = Tregs_list, name = "Tregs", assay = "RNA")
h <- FeaturePlot(object = T_cells, features = "Tregs1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "7_ Tregs", subtitle = "(Cd4, Foxp3,Il2ra, Icos, Tnfrsf4)")+ theme(plot.subtitle = element_text(hjust =0.5))
h
ggsave(file = "figures/Tcells/Tregs_Tregs.png", dpi=300, width=5, height=5)

```



## Annotation


























