---
title: "tils_visualisation"
author: "Michael Monsonego"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    fig_crop: false
    toc: true
    toc_float: true
  pdf_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width=17) 
knitr::opts_chunk$set(cache = TRUE) 
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```
## 1. number of cells per patient + tSNE plot
```{r message=FALSE, warning=FALSE, fig.height= 12, cache=F}


library(dplyr)
library(Seurat)
library(patchwork)
library(tidyverse)
library(sleepwalk)
library(gridExtra)
library(future)
library(reshape2)
library(reticulate)
library(ggplot2)
library(plyr)
library(stringr)
library(RColorBrewer)
library(grid)
library(cowplot)
library(ggsignif)
library(DT)
library(Cairo)
library(snakecase)
library(glue)
library(ggrepel)


setwd("D:/Michael/git_check/tils")
#  setwd("C:/Users/michael monsonego/Documents/tils") #M# update between computers(git)

x1=4.6
#T_cells = readRDS("objects/tils_all_.45_integrgate_annotated (2).rds")
T_cells = readRDS("objects/tils_all_.45_integrgate_annotated_merge_prolif.rds")

# counting cells in treatment and control
unique(T_cells$orig.ident) # R1-4, N1-4

table(T_cells$orig.ident)["R1"] #  
table(T_cells$orig.ident)["R2"] #  
table(T_cells$orig.ident)["R3"] #  
table(T_cells$orig.ident)["R4"] #  
table(T_cells$orig.ident)["N1"] #  
table(T_cells$orig.ident)["N2"] #  
table(T_cells$orig.ident)["N3"] #  
table(T_cells$orig.ident)["N4"] #  
table(T_cells$Treatment)

DimPlot(T_cells, reduction = "tsne", label = TRUE, pt.size = 0.5, label.size = 10) +
  ggplot2::theme(legend.position = "none") +
  ggplot2::theme(
    axis.title.x = element_text(size = 30),  # X-axis title size
    axis.title.y = element_text(size = 30),  # Y-axis title size
    axis.text.x = element_text(size = 24),                 # X-axis tick labels size
    axis.text.y = element_text(size = 24),                 # Y-axis tick labels size
    axis.ticks = element_line(size = 1)                    # Adjust axis tick size
  )
ggsave(file = "figures/Tcells/tsne_no_legend.png", dpi=300, width=15, height=10)




TSNEPlot(T_cells, group.by ="Treatment", cols= c("#A799B7", "#DD614A"), pt.size=0.5)+
  theme(legend.text=element_text(size=20))
ggsave(file = "figures/Tcells/tsne_by_treatment.png", dpi=300, width=8, height=6)
```

Top DE genes
--------------
```{r}
Top50Markers <- read.delim("excels/Top50Markers_perClust.tils.45integrgate.csv", sep = ",", header = T, quote="\"", check.names=FALSE)
Top50Markers$avg_log2FC <- Top50Markers$avg_log2FC %>% round(2)
Top50Markers <- Top50Markers[,c("cluster","gene", "avg_log2FC", "p_val")] 
datatable(Top50Markers, rownames = FALSE, filter = 'top', colnames = c('Cluster #', 'Gene Name', "log2(FC)", 'P.value'))
```

```{r eval=FALSE, fig.height=8, message=FALSE, warning=FALSE, include=FALSE}
## Contribution of each group 
#M# unnormalized counts per cluster per treatment.
ggplot(T_cells@meta.data, aes(x=T_cells@active.ident, fill=Treatment)) + geom_bar(stat="count",position=position_dodge())+
     theme_classic() +  scale_fill_manual(values= c("#A799B7", "#DD614A"))+
    labs(x="")+
    theme(legend.title=element_text(size=18),legend.text= element_text(size=14), 
          axis.text.x=element_text(size=12,angle=45, hjust=1))+labs(title = "unnormalized counts per cluster per treatment")

ggsave(file = "figures/Tcells/unnormalized_counts_per_cluster_per_treatment.png", dpi=300, width=15, height=10)
```

```{r eval=FALSE, fig.height=8, message=FALSE, warning=FALSE, include=FALSE}
#M# percentage of contribution to each cluster per treatment
ggplot(T_cells@meta.data, aes(x=T_cells@active.ident, fill=Treatment)) + geom_bar(position="fill")+
     theme_classic() +  scale_fill_manual(values= c("#A799B7", "#DD614A"))+
    labs(x="")+
    theme(legend.title=element_text(size=18),legend.text= element_text(size=14), 
          axis.text.x=element_text(size=12,angle=45, hjust=1))+labs(title = "percentage of contribution to each cluster per treatment")
ggsave(file = "figures/Tcells/percentage_of_contribution_stacked_bar_plot.png", dpi=300, width=10, height=10)
```

```{r eval=FALSE, fig.height=8, message=FALSE, warning=FALSE, include=FALSE}
## unnormalized contribution per patient 

Sample_Names_vec <- c("N1","N2","N3","N4")
for (sample in Sample_Names_vec) {
  Tsub = T_cells@meta.data[T_cells@meta.data$orig.ident == sample, ]
 ggplot(Tsub, aes(x=Tsub$seurat_clusters, fill=Treatment)) + geom_bar(stat="count",position=position_dodge())+
    scale_fill_manual(values= c("#CFBAE1"))+
     theme_classic()
    labs(x="")+
    theme(legend.title=element_text(size=18),legend.text= element_text(size=14), 
          axis.text.x=element_text(size=12,angle=45, hjust=1))+
      labs(title = glue({sample}, " : unnormalized counts per cluster per treatment"))
  ggsave(filename = glue({sample}, "_unnormalized_counts_per_cluster_per_treatment.png"), path = "figures/Tcells/", dpi=300, width=10, height=10)
}

Sample_Names_vec <- c("R1","R2","R3","R4")
for (sample in Sample_Names_vec) {
  Tsub = T_cells@meta.data[T_cells@meta.data$orig.ident == sample, ]
 ggplot(Tsub, aes(x=Tsub$seurat_clusters, fill=Treatment)) + geom_bar(stat="count",position=position_dodge())+
    scale_fill_manual(values= c("#5C7D9D"))+
     theme_classic() + 
    labs(x="")+
    theme(legend.title=element_text(size=18),legend.text= element_text(size=14), 
          axis.text.x=element_text(size=12,angle=45, hjust=1))+labs(title = glue({sample}, " : unnormalized counts per cluster per treatment"))
  ggsave(filename = glue({sample}, "_unnormalized_counts_per_cluster_per_treatment.png"), path = "figures/Tcells/", dpi=300, width=10, height=10)
}
```
## normalized contribution per patient 
```{r echo=FALSE, fig.height=10, fig.width=10}
frque <- table(Idents(T_cells), T_cells$Treatment)
frque <- as.data.frame(frque)
frque <- frque %>% dplyr::rename(Cluster = Var1, Treatment = Var2)
frque$Treatment <- factor(frque$Treatment)

frque <- ddply(frque, .(Treatment),  transform, percentperstatus   = Freq/sum(Freq))
frque <- ddply(frque, .(Cluster), transform, percentpercluster  = Freq/sum(Freq))
frque <- ddply(frque, .(Cluster), transform, normalizedppstatus = percentperstatus/sum(percentperstatus)*100)
frque <- ddply(frque, .(Treatment),  transform, normalizedpcluster = percentpercluster/sum(percentpercluster))

mfrque <- table(Idents(T_cells), T_cells$orig.ident)
mfrque <- as.data.frame(mfrque)
mfrque <- mfrque %>% dplyr::rename(Cluster = Var1, patient = Var2)

mfrque <- ddply(mfrque, .(patient),  transform, percentperstatus   = Freq/sum(Freq))
mfrque <- ddply(mfrque, .(Cluster), transform, percentpercluster  = Freq/sum(Freq))
mfrque <- ddply(mfrque, .(Cluster), transform, normalizedppstatus = percentperstatus/sum(percentperstatus)*100)
mfrque <- ddply(mfrque, .(patient),  transform, normalizedpcluster = percentpercluster/sum(percentpercluster))


p3 <- ggplot(mfrque, aes(x = Cluster, y = normalizedppstatus, fill = patient)) +
 theme_minimal() +
 geom_bar(stat = "identity", width = 0.7,position="dodge") +
   scale_fill_manual(values= c("#CFBAE1", "#A8A5CA", "#8291B3", "#5C7D9D", "#FF6F61", "#D78459", "#AF9A52", "#88B04B"))+
 ggtitle("Normalized Cluster Counts") +
 ylab("Normalized Proportions in %") +
 theme(
   axis.text.x  = element_text(angle = -40, hjust = 0, size = 30, color = "black"),
   axis.text.y  = element_text(size = 26, color = "black"),
   axis.title.x = element_text(size = 36, color = "black"),
   axis.title.y = element_text(size = 36, color = "black"),
   title = element_text(size = 30, hjust = 0.5),
   legend.justification = "top",
   panel.border = element_blank(),
   legend.text = element_text(size=24)) +
 geom_hline(yintercept = 12.5, linetype='dotted')
  
  
p1 <- ggplot(frque, aes(x = Cluster, y = normalizedppstatus, fill = Treatment)) +
 theme_minimal() +
 geom_bar(stat = "identity", width = 0.7,position="dodge") +
 scale_fill_manual(values= c("#A799B7", "#DD614A"))+
 ggtitle("Normalized Cluster Counts") +
 ylab("Normalized Proportions in %") +
 theme(
   axis.text.x  = element_text(angle = -40, hjust = 0, size = 30, color = "black"),
   axis.text.y  = element_text(size = 26, color = "black"),
   axis.title.x = element_text(size = 36, color = "black"),
   axis.title.y = element_text(size = 36, color = "black"),
   title = element_text(size = 30),
   panel.border = element_blank(),
   legend.text = element_text(size=34)) +
 geom_hline(yintercept = 50, linetype='dotted') 

p2<-ggplot(frque, aes(x= Cluster, y= Freq, fill = Treatment)) +
 theme_minimal() +
 geom_bar(stat = "identity", width = 0.7) +
 scale_fill_manual(values= c("#A799B7", "#DD614A"))+
 ggtitle("Cluster Counts") + ylab("Cell counts\n") +
 theme(
   axis.text.x  = element_text(angle = -40, hjust = 0, size = 30, color = "black"),
   axis.text.y  = element_text(size = 26, color = "black"),
   axis.title.x = element_text(size = 36, color = "black"),
   axis.title.y = element_text(size = 36 , color = "black"),
   title = element_text(size = 30),
   legend.justification = "top",
   panel.border = element_blank(),
   legend.text = element_text(size=34))

p2+p1+plot_layout(guides = "collect")+plot_annotation(tag_levels = 'A') 
ggsave(file = "figures/Tcells/unnormalized_and_normalized_contribution_per_patient.png", dpi=300, width=14, height=10)

p1
ggsave(file = "figures/Tcells/barPlotsSeparated_1.png", dpi=300, width=14, height=10)


p2
ggsave(file = "figures/Tcells/barPlotsSeparated_2.png", dpi=300, width=14, height=10)


p3
ggsave(file = "figures/Tcells/normalized_cluster_counts.png", dpi=300, width=14, height=10)

```

volcano plot
--------------
```{r fig.height=10}
# DE_genes <- FindMarkers(T_cells, ident.1 = "Responder", ident.2 = "Non_Responder", group.by = "Treatment", assay = "RNA")
# DE_genes <- rownames_to_column(DE_genes, var = "gene")
# write_csv(DE_genes, "excels/DE_genes_res_nonRes_all.csv") 
# Top100Markers <- DE_genes %>% 
#   top_n(n = 100, wt = avg_log2FC) %>% 
#   as.data.frame
# write_csv(Top100Markers, "excels/Top100Markers_per_treatment_tils_all.csv") 

DE_genes <- read.csv("excels/DE_genes_res_nonRes_all.csv")

DE_genes$Significance <- "Not Significant"
DE_genes$Significance[DE_genes$p_val_adj < 0.05 & DE_genes$avg_log2FC > 0] <- "Upregulated"
DE_genes$Significance[DE_genes$p_val_adj < 0.05 & DE_genes$avg_log2FC < 0] <- "Downregulated"

genes_of_interest <- c("ILR7", "GNZMB")  # replace with gene names of interest to see on volcano plot

top_1000_genes <- DE_genes %>%
  arrange(p_val_adj) %>%
  head(1000) %>%
  pull(gene) 

label_genes <- union(genes_of_interest, top_1000_genes)
DE_genes$Label <- ifelse(DE_genes$gene %in% label_genes, DE_genes$gene, NA)

ggplot(DE_genes, aes(x = avg_log2FC, y = -log10(p_val_adj))) +
  geom_point(aes(color = Significance)) +
  scale_color_manual(values = c("Not Significant" = "grey", "Upregulated" = "blue", "Downregulated" = "red")) +
  theme_minimal() +
  labs(title = "Volcano Plot", x = "Log2 Fold Change", y = "-Log10 Adjusted P-value") +
  theme(legend.title = element_blank()) +
  geom_text_repel(aes(label = Label), 
                  box.padding = 0.35, 
                  point.padding = 0.3, 
                  max.overlaps = 10, 
                  segment.color = 'grey50')+
  ylim(0, 300)
ggsave(file = "figures/Tcells/volcano.png", dpi=300, width=15, height=10)

```
Top DE genes between responders and non-responders
--------------
```{r}
Top100Markers <- read.delim("excels/Top100Markers_per_treatment_tils_all.csv", sep = ",", header = T, quote="\"", check.names=FALSE)
Top100Markers$avg_log2FC <- Top100Markers$avg_log2FC %>% round(2)
Top100Markers <- Top100Markers[,c("gene", "avg_log2FC", "p_val")] 
datatable(Top100Markers, rownames = FALSE, filter = 'top', colnames = c('Gene Name', "log2(FC)", 'P.value'))
```

Markers
=====

### Proliferation Markers

```{r fig.height=x1*1, cache=TRUE}
Prolif <- c("TOP2A","MKI67")
FeaturePlot(T_cells, features = Prolif, label.size = 8, pt.size = 1, label=F, ncol = 3, order = T,reduction = "tsne")
ggsave(file = "figures/Tcells/Proliferation_Markers.png", dpi=300, width=20, height=x1*2)
```

### T Cells Markers
```{r cache=T, fig.height=x1*2,warning=FALSE}
Ts <- c('Cd4', 'Cd8a', 'Cd3d','Cd3e','Cd3g',"PTPRC")
Ts <- toupper(Ts)
FeaturePlot(T_cells, features = Ts, label.size = 8, pt.size = 1, label=F, ncol = 3, order = T,reduction = "tsne")
ggsave(file = "figures/Tcells/T_Cells_Markers.png", dpi=300, width=15, height=x1*2)
```
### checking 
```{r cache=T, fig.height=x1*3, warning=FALSE}
FeaturePlot(T_cells, features = c("CXCR5", "CXCL13", "BCL6", "ASCL2", "IL21", "PDCD1", "FOXP3"), order=TRUE,pt.size=1, reduction="tsne", ncol=3)
```

```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=20}
VlnPlot(T_cells, features = c("CXCR5", "CXCL13", "BCL6", "ASCL2", "IL21", "PDCD1", "FOXP3"), assay = "RNA", stack=TRUE, flip= TRUE, split.by = "Treatment")+  theme_classic() +  scale_fill_manual(values= c("#A799B7", "#DD614A"))+theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24))+
  theme(
    axis.text.y = element_text(size = 24, face = "italic"),
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.ticks.y = element_line(size = 0.5))+
geom_boxplot(alpha=0.3,show.legend = FALSE)

VlnPlot(T_cells, features = c("CXCR5", "CXCL13", "BCL6", "ASCL2", "IL21", "PDCD1", "FOXP3"), assay = "RNA", stack=TRUE, flip= TRUE, group.by = "Treatment", fill.by ="ident")+theme(axis.text.y =element_text(size=6))+ theme_classic() +
   scale_fill_manual(values= c("#A799B7", "#DD614A"))+
  theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)

VlnPlot(T_cells, features = c("CXCR5", "CXCL13", "BCL6", "ASCL2", "IL21", "PDCD1", "FOXP3"), assay = "RNA", stack=TRUE, flip= TRUE, fill.by = "ident")+theme(axis.text.y =element_text(size=6))+  theme_classic() +theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
```



### a. activation/exhaustion markers
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= 16}
# activation / exhaustion markers
FeaturePlot(T_cells, features = c("PDCD1","HAVCR2","LAG3","TIGIT","TNFRSF4","TNFRSF9","ICOS","CTLA4","KLRD1","IRF8","KLRC1","TOX"), order=TRUE,pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_activation.png", dpi=300, width=10, height=10)
```

```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=20}
VlnPlot(T_cells, features = c("PDCD1","HAVCR2","LAG3","TIGIT","TNFRSF4","TNFRSF9"), assay = "RNA", stack=TRUE, flip= TRUE, split.by = "Treatment")+  theme_classic() +  scale_fill_manual(values= c("#A799B7", "#DD614A"))+theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) +
   scale_fill_manual(values= c("#A799B7", "#DD614A"))+
  theme(
    axis.text.y = element_text(size = 24, face = "italic"),
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.ticks.y = element_line(size = 0.5))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
 ggsave(filename = "vln_activation_exhastion_cluster_by_treatment_1.png" , path = "figures/Tcells/", dpi=300, width=12, height=10)

VlnPlot(T_cells, features = c("PDCD1","HAVCR2","LAG3","TIGIT","TNFRSF4","TNFRSF9"), assay = "RNA", stack=TRUE, flip= TRUE, group.by = "Treatment", fill.by ="ident")+theme(axis.text.y =element_text(size=6))+  theme_classic() +  scale_fill_manual(values= c("#A799B7", "#DD614A"))+theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/vln_activation_exhastion_by_treatment_1.png", dpi=300, width=6, height=10)

VlnPlot(T_cells, features = c("PDCD1","HAVCR2","LAG3","TIGIT","TNFRSF4","TNFRSF9"), assay = "RNA", stack=TRUE, flip= TRUE, fill.by = "ident")+theme(axis.text.y =element_text(size=6))+  theme_classic() +theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/vln_activation_exhastion_by_cluster_1.png", dpi=300, width=6, height=10)

##

VlnPlot(T_cells, features = c("ICOS","CTLA4","KLRD1","IRF8","KLRC1","TOX"), assay = "RNA", stack=TRUE, flip= TRUE, split.by = "Treatment")+theme(axis.text.y =element_text(size=4))+  theme_classic() +  scale_fill_manual(values= c("#A799B7", "#DD614A"))+theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_activation_exhaustion_cluster_by_treatment_2.png", dpi=300, width=8, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("ICOS","CTLA4","KLRD1","IRF8","KLRC1","TOX"), assay = "RNA", stack=TRUE, flip= TRUE, group.by = "Treatment", fill.by ="ident")+theme(axis.text.y =element_text(size=4))+  theme_classic() +  scale_fill_manual(values= c("#A799B7", "#DD614A"))+theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_activation_exhaustion_by_treatment_2.png", dpi=300, width=4, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("ICOS","CTLA4","KLRD1","IRF8","KLRC1","TOX"), assay = "RNA", stack=TRUE, flip= TRUE, fill.by = "ident")+theme(axis.text.y =element_text(size=4))+  theme_classic() +
  theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_activation_exhaustion_by_cluster_2.png", dpi=300, width=6, height=10, limitsize=FALSE)

```

### b. effector markers 
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=16}
FeaturePlot(T_cells, features = c("GZMB","PRF1","IFNG","MKI67","GZMC","GZMA","KLRG1","IL7R","CD52"),order=TRUE,pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_effector.png", dpi=300, width=8, height=10)
```
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=20}
VlnPlot(T_cells, features = c("GZMB","PRF1","IFNG","TNF"), assay = "RNA", stack=TRUE, flip= TRUE, split.by = "Treatment")+theme(axis.text.y =element_text(size=6))+  theme_classic() +  scale_fill_manual(values= c("#A799B7", "#DD614A"))+theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12)) +
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_effector_cluster_by_treatment_1.png", dpi=300, width=12, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("GZMB","PRF1","IFNG","TNF"), assay = "RNA", stack=TRUE, flip= TRUE, group.by = "Treatment", fill.by ="ident")+theme(axis.text.y =element_text(size=6))+  theme_classic() +  scale_fill_manual(values= c("#A799B7", "#DD614A"))+theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12)) +
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_effector_by_treatment_1.png", dpi=300, width=6, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("GZMB","PRF1","IFNG","TNF"), assay = "RNA", stack=TRUE, flip= TRUE, fill.by = "ident")+theme(axis.text.y =element_text(size=6))+  theme_classic() +  theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12)) +
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_effector_by_cluster_1.png", dpi=300, width=6, height=10, limitsize=FALSE)

##

VlnPlot(T_cells, features = c("GZMC","GZMA","KLRG1","CD52"), assay = "RNA", stack=TRUE, flip= TRUE, split.by = "Treatment")+theme(axis.text.y =element_text(size=6))+  theme_classic()+  scale_fill_manual(values= c("#A799B7", "#DD614A")) +theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_effector_cluster_by_treatment_2.png", dpi=300, width=8, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("GZMC","GZMA","KLRG1","CD52"), assay = "RNA", stack=TRUE, flip= TRUE, group.by = "Treatment", fill.by ="ident")+
  theme(axis.text.y =element_text(size=6))+
  theme_classic() +
  scale_fill_manual(values= c("#A799B7", "#DD614A"))+
  theme(
    axis.text.x = element_text(angle = 50, hjust=1, size=12),
    axis.title.x = element_blank(),
    legend.text = element_text(size=24),
    axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_effector_by_treatment_2.png", dpi=300, width=4, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("GZMC","GZMA","KLRG1","CD52", fill.by = "ident"), assay = "RNA", stack=TRUE, flip= TRUE)+theme(axis.text.y =element_text(size=6))+  theme_classic() +theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_T_effector_by_cluster_2.png", dpi=300, width=6, height=10, limitsize=FALSE)

```

### CD69 + CD39
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= x1, fig.width=8}
# naive
FeaturePlot(T_cells, features = c("CD69","ENTPD1"),order=TRUE,pt.size=1, reduction="tsne", ncol=2)
ggsave(file = "figures/Tcells/CD69.png", dpi=300, width=8, height=10)
```
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= x1}


VlnPlot(T_cells, features = c("CD69","ENTPD1"), fill.by = "ident", assay = "RNA", stack=TRUE, flip= TRUE)+theme(axis.text.y =element_text(size=6))+  theme_classic() +theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
ggsave(file = "figures/Tcells/Vln_CD69.png", dpi=300, width=6, height=10, limitsize=FALSE)

VlnPlot(T_cells, features = c("CD69"), assay = "RNA", flip= TRUE, split.by = "Treatment", pt.size = 0)+theme(axis.text.y =element_text(size=6))+  theme_classic() +  scale_fill_manual(values= c("#A799B7", "#DD614A"))+theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)

VlnPlot(T_cells, features = c("ENTPD1"), assay = "RNA", flip= TRUE, split.by = "Treatment", pt.size = 0)+theme(axis.text.y =element_text(size=6))+  theme_classic() +  scale_fill_manual(values= c("#A799B7", "#DD614A"))+theme(axis.text.x = element_text(angle = 50, hjust=1, size=12), axis.title.x = element_blank(), legend.text = element_text(size=24)) + theme(axis.title.y=element_text(size=12))+
geom_boxplot(alpha=0.3,show.legend = FALSE)
```


### c. naive-like (naive or memory cells)
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= x1}
# naive
FeaturePlot(T_cells, features = c("SELL","IL7R","CCR7","MS4A4C"),order=TRUE,pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_naive.png", dpi=300, width=8, height=10)
```

```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height=x1*2}
# memory 
FeaturePlot(T_cells, features = c("SELL","CD44","IL7R","CCR7","TCF7","CXCR6"),order=TRUE,pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_memory.png", dpi=300, width=10, height=8)
```

### d. effector memory markers
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= x1*4}
FeaturePlot(T_cells, features = c("BACH2","PRDM1","IL7R","CD44","CX3CR1","ID2","KLF2","KLRG1","LEF1","STAT4","TBX21","ZEB2"),order=TRUE,pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_effector memory.png", dpi=300, width=30, height=40)
```

### e. central memory markers
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= x1*5}
FeaturePlot(T_cells, features = c("TCF7","BACH2","BCL16","CCR7","IL7R","CD27","CD44","SELL","ID3","KLF2","LCR7","LEF1","SLAMF6","STAT3","TFAP4"),order=TRUE,pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_central memory.png", dpi=300, width=30, height=50, limitsize=FALSE)
```


### f. resident memory markers
```{r message=FALSE, warning=FALSE, cache= TRUE, fig.height= x1*8}
FeaturePlot(T_cells, features = c("PRDM1","ITGAE","IL7R","CD44","ITGA1","CD69","CRTAM","DUSP6","EGR2","FOSB","FOXO3","GZMA","GZMB","ZNF683","ID3","IL2RA","IRF4","JUN","JUNB","MYC","NR4A1","RGS1"), order=TRUE, pt.size=1, reduction="tsne", ncol=3)
ggsave(file = "figures/Tcells/features_T_resident memory.png", dpi=300, width=30, height=80, limitsize=FALSE)
```

# add signatures to final html, takes time and cashe = FALSE manditory
### h. T cells features signatures 
```{r eval=FALSE, fig.height=4, fig.width=4, message=FALSE, warning=FALSE, cache=FALSE, include=FALSE}
CD8_memory_list = list(c("IL7R","CD44","CXCR6"))
T_cells = AddModuleScore(object = T_cells, features = CD8_memory_list, name = "CD8_memory", assay = "RNA")
a <- FeaturePlot(object = T_cells, features = "CD8_memory1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "0_ CD8+ memory", subtitle = "(IL7R,CD44,CXCR6)")+ theme(plot.subtitle = element_text(hjust = 0.5))
a
ggsave(file = "figures/Tcells/CD8_memory_Cxcl10.png", dpi=300, width=5, height=5)

activated_list = list(c("PDCD1","KLRD1","GZMA"))
T_cells = AddModuleScore(object = T_cells, features = activated_list, name = "activated", assay = "RNA")
b <- FeaturePlot(object = T_cells, features = "activated1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "activated_CD8+activated", subtitle = "(PDCD1,KLRD1,GZMA)")+ theme(plot.subtitle = element_text(hjust =0.5))
b
ggsave(file = "figures/Tcells/activated.png", dpi=300, width=5, height=5)

effector_memory_list = list(c("ID2","STAT3","GZMB","PRF1","IL7R"))
T_cells = AddModuleScore(object = T_cells, features = effector_memory_list, name = "effector_memory", assay = "RNA")
c <- FeaturePlot(object = T_cells, features = "effector_memory1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "2_ CD8+ effector memory", subtitle = "(ID2,STAT3,GZMB,PRF1,IL7R)")+ theme(plot.subtitle = element_text(hjust =0.5))
c
ggsave(file = "figures/Tcells/effector_memory_Tregs.png", dpi=300, width=5, height=5)

cytotoxic_list = list(c("GZMB","GZMC","GRZMF", "IFNG","PRF1", "TNF"))
T_cells = AddModuleScore(object = T_cells, features = cytotoxic_list, name = "cytotoxic", assay = "RNA")
d <- FeaturePlot(object = T_cells, features = "cytotoxic1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "3_ CD8+ cytotoxic", subtitle = "(GZMB,GZMC,GRZMF)")+ theme(plot.subtitle = element_text(hjust =0.5))
d
ggsave(file = "figures/Tcells/cytotoxic_Tregs.png", dpi=300, width=5, height=5)


CD8_naive_memory_list = list(c("CD8A","CD4","SELL","CCR7","TCF7","LEF1","CD8","ITGAE","CXCR3","GZMA","SELL","CCR7"))
T_cells = AddModuleScore(object = T_cells, features = CD8_naive_memory_list, name = "CD8_naive_memory", assay = "RNA")
e <- FeaturePlot(object = T_cells, features = "CD8_naive_memory1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "4_ CD8+ naive/memory", subtitle = "(CD8A,CD4,SELL,CCR7,TCF7,LEF1,CD8,ITGAE,CXCR3,GZMA,SELL,CCR7)")+ theme(plot.subtitle = element_text(hjust =0.5))
e
ggsave(file = "figures/Tcells/CD8_naive_memory_Tregs.png", dpi=300, width=5, height=5)


CD4_CD8_naive_memory_list = list(c("CD8A","CD4","SELL","CCR7","TCF7","LEF1"))
T_cells = AddModuleScore(object = T_cells, features = CD4_CD8_naive_memory_list, name = "CD4_CD8_naive_memory", assay = "RNA")
f <- FeaturePlot(object = T_cells, features = "CD4_CD8_naive_memory1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "5_ CD4+ CD8+ naive/memory", subtitle = "(CD8A,CD4,SELL,CCR7,TCF7,LEF1)")+ theme(plot.subtitle = element_text(hjust =0.5))
f
ggsave(file = "figures/Tcells/CD4_CD8_naive_memory_CD4_CD8_naive_memory.png", dpi=300, width=5, height=5)

proliferating_list = list(c("CD8","MKI67"))
T_cells = AddModuleScore(object = T_cells, features = proliferating_list, name = "proliferating", assay = "RNA")
g <- FeaturePlot(object = T_cells, features = "proliferating1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "6_ CD8+ proliferating", subtitle = "(Cd8, Mki67)")+ theme(plot.subtitle = element_text(hjust =0.5))
g
ggsave(file = "figures/Tcells/proliferating_CD8+proliferating.png", dpi=300, width=5, height=5)

Tregs_list = list(c("CD4","FOXP3","IL2RA","ICOS","TNFRSF4"))
T_cells = AddModuleScore(object = T_cells, features = Tregs_list, name = "Tregs", assay = "RNA")
h <- FeaturePlot(object = T_cells, features = "Tregs1", reduction = "tsne", cols=c("grey","grey","#e46467", "#b33336", "#A73033"))+labs(title = "7_ Tregs", subtitle = "(Cd4, Foxp3,Il2ra, Icos, Tnfrsf4)")+ theme(plot.subtitle = element_text(hjust =0.5))
h
ggsave(file = "figures/Tcells/Tregs_Tregs.png", dpi=300, width=5, height=5)
```



```{r eval=FALSE, fig.height=4, fig.width=7, cache=FALSE, include=FALSE}
## Annotation

Idents(T_cells) <- 'seurat_clusters'
T_cells <- RenameIdents(T_cells, "0" = "0_CD4_activated_Tmem",
                                 "1" = "1_cytotoxic_CD8",
                                 "2" = "2_cytotoxic_CD8",
                                 "3" = "3_CD8_DNA_replication",
                                 "4" = "4_CD8_h.s_proteins",
                                 "5" = "5_CD8_cell_cycle",
                                 "6" = "6_CD8_exhausted",
                                 "7" = "7_CD8_proliferation_like",
                                 "8" = "8_CD8_proliferation")

p <- DimPlot(T_cells, reduction = "tsne", label = F, label.size = 3.5)
LabelClusters(p, id = "ident",  fontface = "bold",position = "median")
#M# saveRDS(T_cells, file = "objects/tils_all_.45_integrgate_annotated.rds")

# merge proliferation clusters(5,7,8) in all cells--------------------------------------------

new.cluster.ids <- c("0" = "0_CD4_activated_Tmem",
                     "1" = "1_cytotoxic_CD8",
                     "2" = "2_cytotoxic_CD8",
                     "3" = "3_CD8_DNA_replication",
                     "4" = "4_CD8_h.s_proteins",
                     "5" = "5_CD8_proliferation",
                     "6" = "6_CD8_exhausted",
                     "7" = "5_CD8_proliferation",
                     "8" = "5_CD8_proliferation")
names(new.cluster.ids) <- levels(T_cells)
T_cells <- RenameIdents(T_cells, new.cluster.ids)
DimPlot(T_cells, reduction = "tsne", label = TRUE, pt.size = 0.5) + NoLegend()


#M# saveRDS(T_cells, file = "objects/tils_all_.45_integrgate_annotated_merge_prolif.rds")
```




&nbsp;
<hr />
<p style="text-align: center;">Work by Michael Monsonego</a></p>
<p style="text-align: center;"> <a href="https://www.asafmadilab.com/">Madi Lab</a></p>
&nbsp;





















